Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-08-30 paper/ms.tex
Small manuscript changes

% Abstract of the paper
\begin{abstract}
Elongated bar-like features are ubiquitous in galaxies, occurring at the centers
of approximately two-thirds of spiral disks.  Due to gravitational interactions
between the bar and the other components of galaxies, it is expected that
angular momentum and matter will redistribute over long (Gyr) timescales in
barred galaxies. Previous work ignoring the gas phase of galaxies has
conclusively demonstrated that bars should slow their rotation over time due to
their interaction with dark matter halos. We have performed a simulation of a
Milky Way-like galactic disk hosting a strong bar which includes a
state-of-the-art model of the interstellar medium and a live dark matter halo.
In this simulation the bar pattern does not slow down over time, and instead
remains at a stable, constant rate of rotation. This behavior has been observed
in previous simulations using more simplified models for the interstellar gas
but it has remained unexplained. We propose that the gas phase of the disk and
the dark matter halo act in concert to stabilize the bar pattern speed and
prevent the bar from slowing down or speeding up. We find that in a Milky
Way-like disk, a gas fraction of only about 5\% is necessary for this mechanism
to operate. Our result naturally explains why nearly all observed bars rotate
rapidly and is especially relevant for our understanding of how the Milky Way
arrived at its present state.
\end{abstract}

@@

Approximately two-thirds of spiral disks host an elongated bar-like feature at
their centers \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our
own Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of
bars is not difficult to explain, since stellar disks simulated in isolation
almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
studies have shown that a spherical potential, such as a stellar bulge or dark
matter halo, acts to stabilize a stellar disk against bar formtaion
\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.

>>>>

% Abstract of the paper
\begin{abstract}
Elongated bar-like features are ubiquitous in galaxies, occurring at the centers
of approximately two-thirds of spiral disks.  Due to gravitational interactions
between the bar and the other components of galaxies, it is expected that
angular momentum and matter will redistribute over long (Gyr) timescales in
barred galaxies. Previous work ignoring the gas phase of galaxies has
conclusively demonstrated that bars should slow their rotation over time due to
their interaction with dark matter halos. We have performed a simulation of a
Milky Way-like galactic disk hosting a strong bar which includes a
state-of-the-art model of the interstellar medium and a live dark matter halo.
In this simulation the bar pattern does not slow down over time, and instead
remains at a stable, constant rate of rotation. This behavior has been observed
in previous simulations using more simplified models for the interstellar gas
but the apparent lack of secular evolution has remained unexplained. We propose
that the gas phase of the disk and the dark matter halo act in concert to
stabilize the bar pattern speed and prevent the bar from slowing down or
speeding up. We find that in a Milky Way-like disk, a gas fraction of only about
5\% is necessary for this mechanism to operate. Our result naturally explains
why nearly all observed bars rotate rapidly and is especially relevant for our
understanding of how the Milky Way arrived at its present state.
\end{abstract}

@@

Approximately two-thirds of spiral disks host an elongated bar-like feature at
their centers \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our
own Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of
bars is not difficult to explain, since stellar disks simulated in isolation
almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
studies have shown that a spherical potential, such as a stellar bulge or dark
matter halo, acts to stabilize stellar disks against bar formtaion
\citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.
